\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2021-10-07 lin-vs-cen.tex
Change of wording

A recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).  If we take $\tw(G)=k^{38}$ and $b=k$ then, for any graph $G$ with $\td(G)\ge k^{39}$, one of the following cases applies:
\begin{enumerate}
    \item $\tw(G)\ge k^{38}$: In this case the proof follows Steps~2 and 3 as before to show that $\chilin(G)\in\Omega(k)$.
    \item $G$ contains a tree $T$ of treedepth $k$.  The same paper \cite[Lemma~1.6]{czerwinski.nadara.ea:improved} shows that any tree $T$ of treedepth $k$ contains a subcubic subtree of treedepth $\Omega(k)$.
    A theorem of \citet{kun.obrien.ea:polynomial} shows that, for any subcubic tree $T$, $\chicen(T) \le (\log_2 3)\cdot\chilin(T)$.  Therefore, in this case $\chilin(G)\ge\chilin(T)\ge \chicen(T)/\log_2 3 = \Omega(k)$.
\end{enumerate}

@@

As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.  This, by itself, reduces the exponent in \cref{kun_obrien_general2} from $39$ to $20$.  We observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.  For example, if we consider only planar graphs, then it is well known that any planar graph $G$ of treewidth $k$ contains an $\Omega(k\times k)$ grid minor.  This already reduces the exponent in \cref{kun_obrien_general2} further (for planar graphs) to $3$ ($2$ if we can prove the result for grid minors).  If one could prove a similar result for the first ``heavy hammer'' then this would show that, for any planar graph $G$, $\chicen(G)\in O(\chilin(G))$.

@@

\begin{proof}
  I have no idea how to prove this.  Computer search?.
\end{proof}

@@

\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}

>>>>

A recent result of \citet[Theorem~1.4]{czerwinski.nadara.ea:improved} shows that if $b$ is the maximum treedepth of any tree contained in $G$ then $\td(G)\le b\cdot\tw(G)$ (and the proof is easy).  Therefore, if $\td(G)\ge k^39$ then at least one of the following cases applies:
\begin{enumerate}
    \item $\tw(G)\ge k^{38}$: In this case the proof follows Steps~2 and 3 as before to show that $\chilin(G)\in\Omega(k)$.
    \item $G$ contains a tree $T$ of treedepth $k$.  The same paper \cite[Lemma~1.6]{czerwinski.nadara.ea:improved} shows that any tree $T$ of treedepth $k$ contains a subcubic subtree of treedepth $\Omega(k)$.
    A theorem of \citet{kun.obrien.ea:polynomial} shows that, for any subcubic tree $T$, $\chicen(T) \le (\log_2 3)\cdot\chilin(T)$.  Therefore, in this case $\chilin(G)\ge\chilin(T)\ge \chicen(T)/\log_2 3 = \Omega(k)$.
\end{enumerate}

@@

As a next step, we would like to improve the third part of the argument to show that, if $G$ contains a $k\times k$ grid minor, then $\chilin(G)\in \Omega(k)$.  This, by itself, would reduce the exponent in \cref{kun_obrien_general2} from $39$ to $20$.  We observe that the first two parts of the argument are both ``heavy hammers'' and that, for some special graph classes, much lighter hammers suffice.  For example, if we consider only planar graphs, then it is well known that any planar graph $G$ of treewidth $k$ contains an $\Omega(k\times k)$ grid minor.  This already reduces the exponent in \cref{kun_obrien_general2} further (for planar graphs) to $3$ ($2$ if we can prove the result for grid minors).  If one could prove a similar result for the first ``heavy hammer'' then this would show that, for any planar graph $G$, $\chicen(G)\in O(\chilin(G))$.

@@

\begin{proof}
  Case analysis(?).
\end{proof}

@@

\todo[inline]{We can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}
